---
title: "TEOS HydroStation S Data Lab"
author: "Irais Luquis Ramos"
date: "2023-03-07"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

#Load required libraries

```{r message=FALSE}
library(tidyverse)
library(gsw)
library(ggplot2)
#install.packages('seacarb')
#install.packages('prettydoc')
```

#Now we need to import our data

```{r warning=FALSE, message=FALSE}
library(readr)
hydrostation_bottle <- read_table("hydrostation_bottle.txt", 
    col_names = FALSE, skip = 31)

library(readr)
hydrostation_bottle_names <- read_csv("hydrostation_bottle.txt", 
    skip = 30)


colnames(hydrostation_bottle)=colnames(hydrostation_bottle_names)
#view(hydrostation_bottle)
```

Variable Names and Units

-   yyyymmdd = Year Month Day\
-   decy = Decimal Year\
-   time = Time (hhmm)\
-   latN = Latitude (Deg N)
-   lonW = Longitude (Deg W)
-   Depth = Depth (m)\
-   Temp = Temperature ITS-90 (C)
-   Pres = CTD Pressure (dbar)\
-   CTD_S = CTD Salinity (PSS-78)\
-   Sal1 = Salinity-1 (PSS-78)\
-   Sig-th = Sigma-Theta (kg/m\^3)\
-   O2(1) = Oxygen-1 (umol/kg)\
-   OxFixT = Oxygen Fix Temp (C)\
-   Anom1 = Oxy Anomaly-1 (umol/kg)\
    Quality flags
-   -999 = No data
-   0 = Less than detection limit

```{r}
#lets firts plot the data
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x = decy, y = `Sig-th`))


hydrostation_bottle %>%
  filter(`Sig-th`!=-999) %>%
  ggplot()+
  geom_point(aes(x = decy, y = `Sig-th`))

hydrostation_bottle %>%
  filter(`Sig-th`!=-999 & Depth <20) %>%
  ggplot()+
  geom_point(aes(x = decy, y = `Sig-th`))

hydrostation_bottle %>%
  filter(`Sig-th`!=-999 & Depth <20) %>%
  ggplot()+
  geom_line(aes(x = decy, y = `Sig-th`))

#clear seasonal signal for sigma-theta,lets see how this compares to temperature
hydrostation_bottle %>%
  filter(`Sig-th`!=-999 & Depth <20) %>%
  ggplot()+
  geom_point(aes(x = Temp, y = `Sig-th`))
#Temperatures and density ares strongly correleted, but there appears to be two outliers tht we sill likley need to adreess at some point
#we only have density data from 1988-present, but temperature and salinity data from 1950s-present 
#This means I can calculate seawater density from 1950s-present 
#we can use the TEOS-10 to do this



```
## TEOS-10 Toolbox in Package seacarb

```{r}
?gsw #launches the documentation for the gibbs seawater toolbox (TEOS-10). All of this function within the package 

?gsw_sigma0 #lets check this function 
#it says we need absolute salinity and conservative temperature USAGE

#first we need absolute salinity:
?gsw_SA_from_SP
#practical salinity
#sea pressure (dbar)
#longitude 
#latitude

#Lets plot our pressure data - its missing before 1980s
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy, y=Pres))

#we have depth data for the time series 
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy, y=Depth))

?gsw_p_from_z

#adds a pressure colum from the depth and latN colums from/to hydrostation bottle #creates new variable

hydrostation_bottle=
hydrostation_bottle %>%
  mutate(Pres_gsw=gsw_p_from_z(Depth*-1,latN))

hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=Pres, y=Pres_gsw))
#we see strong 1:1 agreement between measured pressure and calculated pressure

hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy, y=latN))

hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy, y=lonW))

hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy, y=CTD_S))

#Checking lat, lon and salinity data
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy, y=Sal1))

hydrostation_bottle=
hydrostation_bottle %>%
  mutate(Pres_gsw=gsw_p_from_z(Depth*-1,latN)) %>%
  mutate(S_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN))

#chek it 
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy, y=S_abs_gsw))
#how else can I check my data
hydrostation_bottle %>%
  filter(Sal1!=-999) %>%
  ggplot()+
  geom_point(aes(x=Sal1, y=S_abs_gsw))

#now we need to calculate conservative temperature
?gws_CT_from_t
#we need absolute salinity, in-situ temp (ITS-90), and sea pressure 

#add line to calculate conservative temperature and change from hydrostation_bottle to HydrosS to not create confusion
HydroS=
hydrostation_bottle %>%
  filter(Sal1!=-999) %>% #filter for salinity missing
  filter(Temp!=-999) %>%
  mutate(Pres_gsw=gsw_p_from_z(Depth*-1,latN)) %>% #pressure for depth
  mutate(S_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN)) %>%
  mutate(T_cons_gsw=gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw))

#missing data por eso se añadio filter 

#hydrostation_bottle %>%
  #filter(Sal1!=-999) %>%
  #ggplot()+
  #geom_point(aes(x=Sal1, y=S_abs_gsw))

#hydrostation_bottle %>%
  #filter(Temp!=-999) %>%
  #ggplot()+
  #geom_point(aes(x=Temp, y=T_cons_gsw))

HydroS=
hydrostation_bottle %>%
  filter(Sal1!=-999) %>% 
  filter(Temp!=-999) %>%
  mutate(Pres_gsw=gsw_p_from_z(Depth*-1,latN)) %>% 
  mutate(S_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN)) %>%
  mutate(T_cons_gsw=gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw)) %>%
  mutate(sig_th_gsw=gsw_sigma0(S_abs_gsw,T_cons_gsw))
  
HydroS %>%
  filter(`Sig-th` !=-999) %>%
  ggplot()+
  geom_point(aes(x=`Sig-th`,y=sig_th_gsw))
#but we have a very low sig-th-gsw so lets find it 

#our bottle and ctd salonity do not agree, this is likley the problem for the low
HydroS %>%
  filter(sig_th_gsw<0) %>%
  view()

HydroS_correctedS_a=
  HydroS %>% 
  filter(sig_th_gsw<0) %>% #filter for negative value 
  mutate(S_abs_gsw=gsw_SA_from_SP(CTD_S,Pres_gsw,360-lonW,latN)) %>%
  mutate(T_cons_gsw=gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw)) %>%
  mutate(sig_th_gsw=gsw_sigma0(S_abs_gsw,T_cons_gsw)) 

HydroS_correctedS_b=
  HydroS %>% 
  filter(sig_th_gsw>0) 

HydroS_correctedS = rbind(HydroS_correctedS_a,HydroS_correctedS_b)

HydroS_correctedS %>% 
  filter(`Sig-th`!=-999) %>%
  ggplot()+
  geom_point(aes(x=`Sig-th`,y=sig_th_gsw))

#




  
```


